Optimal Light Curve Attitude Inversion with Measurement Noise: Two Case Studies

### 2025-03-20 paper.tex
added discussion of new results plots

@@

If no initial attitude guess is available, as in most cases of passive debris observation, the inversion problem must be solved globally. Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work, but still rely on high-quality priors as with the Kalman filters. In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024} can be applied to search the solution space exhaustively.

@@

In this paper, we describe a global full-state attitude solver inspired by recent PSO work \cite{burton2024journal} but is easier to tune, can be fully parallelized, and is robust to significant measurement noise \cite{robinson2025att}. Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve. Our approach was designed specifically for use with real data as we take time-varying measurement noise and uncertainties in the body moments of inertia into account explicitly.

@@

Our multi-start solution method begins by creating $n$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge. In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:

\begin{equation}
 \vctr{x}_{0,i}(0) = \begin{bmatrix}\vctr{p}_{i,0}(0) & \vctr{\omega}_{i,0}(0) & \left(J_y / J_x\right)_{i,0} & \left(J_z / J_x\right)_{i,0}\end{bmatrix}^T.
\end{equation}

\noindent
In this work, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is uniformly randomly sampled from the ball in $\mathbb{R}^3$ of radius $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize guesses -- while $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ are sampled from Gaussian prior distributions centered at estimated values derived from the known geometry of the object. We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:

@@

@@

@@

\noindent
producing a final initial state which is a good local guess for a state that produces the observations. In this way, we globally search the solution space up to the density implied by $n$. Using a simpler algorithm like gradient descent can still reach convergence, but will take significantly more objective function evaluations due to the high nonlinearity of the attitude inversion problem.

@@

\begin{equation}
  e^{-f(\vctr(x)_{i,k}(0))} > \ell e^{-f(\vctr(x)^\star_{k}(0))}.
\end{equation}

Here, $\vctr(x)^\star_{k}(0) = \argmin_{\vctr{x}_{i,k}(0)} f(\vctr(x)_{i,k}(0)) $ is the highest likelihood solution found after optimization.

@@

@@

As the objects we investigate in this work are well-approximated with relatively simple non-convex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection. Further explanation of the polygon clipping procedure for computing the total shadowed area and accelerating this computation during inversion with the so-called shadow cache is included here \textbf{TODO: should this be a here? wording this felt weird if we're avoiding saying like "the authors'"} \cite{robinson2025att}.

@@

\section{Attitude Inversion Results for Synthetic Data}

@@

To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth is known. In order to make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used in inversion. We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}.

@@

\begin{table}[H]
  \centering
  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters were estimated qualitatively}
  \vspace*{6pt}
  \begin{tabular}{|l|l|l|l|}
  \hline
  \textbf{Material} & $C_d$ & $C_s$ & $n$ \\ \hline
 Aluminum \cite{fankhauser2023} & $0.34$ & $0.40$ & $8.9$ \\ \hline
 White paint & $0.9$ & $0.1$ & $1$ \\ \hline
  \end{tabular}
  \label{tb:synth_matprops}
\end{table}

@@

\begin{table}[H]
  \centering
  \caption{True object attitude profile for all test cases}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
 Initial $\vctr{p}(0)$ & $-\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
 True $\vctr{\omega}(0)$ ($\text{rad}/\text{s}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
 Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
 External torque $\vctr{M}$ & $\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
  \end{tabular}
  \label{tb:synth_att}
\end{table}
\FloatBarrier

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\figmed]{light_curves_mag_rocket_body_sdc.png}
  \caption{Simulated mean and sampled observations for the rocket body in V-band magnitude, with the variance included in blue.}
  \label{fig:obs_mag_synth}
\end{figure}

@@

\subsection{Synthetic Light Curve Optimization Results}

@@

A total of $10^5$ initial state guesses are created via the sampling scheme described in Section \ref{sec:run_solver} with $\norm{\vctr{\omega}}_\text{max}=5.26$ degrees per second. The inertia tensor ratios are initialized by sampling a symmetric Gaussian distribution with $\sigma=0.05 \: [kg\cdot m^2]$ on each axis to simulate a moderate level of inertia uncertainty. BFGS minimization is then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}. Classifying candidate solutions with $\ell = 1/10$ produces over $10^4$ solutions within this bound. Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions. There is good agreement between the candidate light curves and the observations, indicating that high-quality local minima have been identified.

\begin{figure}[H]
  \centering
  \includegraphics[width=\figmed]{converged_solutions_1_Rocket Body.png}
  \caption{Candidate solution light curves found through BFGS optimization.}
  \label{fig:conv_lcs_synth}
\end{figure}

Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity vectors of the best solutions determined via the procedure in Section \ref{sec:candidate_sols}. We notice that for both the initial orientation in Figure \ref{fig:sigma_sols1} and angular velocity in Figure \ref{fig:omega_sols1}, the ground truth, plotted in red, lies squarely on a candidate solution surface. In angular velocity space, shown in Figure \ref{fig:omega_sols1}, most solutions fall onto the surface of two open-ended cylinders -- one with half the radius of the larger cylinder containing the truth. Similarly in orientation space, shown in Figure \ref{fig:sigma_sols1}, the solutions fall on two distinct manifolds -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}. One of these orientation-space surfaces contains the ground truth. These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce nearly identical observations.

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sigma_solsRocket Body.png}
    \caption{Uniform prior}
    \label{fig:sigma_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sigma_solsRocket Body Prior.png}
    \caption{Informative prior}
    \label{fig:sigma_sols1b}
  \end{subfigure}

  \caption{Candidate solution initial orientation MRP vectors, ground truth highlighted in red.}
  \label{fig:sigma_sols1}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{omega_sols_Rocket Body.png}
    \caption{Uniform prior}
    \label{fig:omega_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{omega_sols_Rocket Body Prior.png}
    \caption{Informative prior}
    \label{fig:omega_sols1b}
  \end{subfigure}

  \caption{Candidate solution initial body-frame angular velocity vectors, ground truth highlighted in red. TODO: change the perspective on the angular velocities}
  \label{fig:omega_sols1}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body.png}
    \caption{Uniform prior}
    \label{fig:i_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body Prior.png}
    \caption{Informative prior}
    \label{fig:i_sols1b}
  \end{subfigure}

  \caption{Candidate solution inertia ratios, ground truth highlighted in red.}
  \label{fig:i_sols1}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body.png}
    \caption{Uniform prior}
    \label{fig:w_vs_ang_error_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body Prior.png}
    \caption{Informative prior}
    \label{fig:w_vs_ang_error_sols1b}
  \end{subfigure}

  \caption{Candidate solution angular orientation error and angular velocity magnitudes compared to the truth}
  \label{fig:w_vs_ang_error_sols1}
\end{figure}

@@

% The inversion

\section{Attitude Inversion Results for Real Data}

@@

\subsection{Delta I Upper Stage}

The Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}. It is $d=0.93$ meters in diameter and is approximately $h=1.7$ meters in length \cite{star37e_astronautix, star37_gunter}.

\begin{figure}[H]
  \centering
  \includegraphics[width=\figsmall]{star37e.JPG}
  \caption{Star-37E upper stage \cite{star37_af}.}
  \label{fig:star37e}
\end{figure}

@@

This light curve shows clear periodicity, but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time. Inversion was performed using $10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}. The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.

\begin{table}[H]
  \centering
  \caption{Delta I rocket body inversion assumptions}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
 Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \hline
 $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} / \text{s}^{-1}$ \\ \hline
 $\ell$ & $1/100$ \\ \hline
\end{tabular}
  \label{tb:case2_ass}
\end{table}

\begin{figure*}[t]
  \centering
  \includegraphics[width=0.8\textwidth]{sphx_glr_results_002_2_00x.png}
  \caption{Candidate solution light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
  \label{fig:case2_s}
\end{figure*}

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sphx_glr_results_006.png}
    \caption{MRP}
    \label{fig:case2_pwa}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sphx_glr_results_005.png}
    \caption{Angular velocity}
    \label{fig:case2_pwb}
  \end{subfigure}

  \caption{Candidate solution orientation MRPs and angular velocity vectors for the Delta I (Star-37E) rocket body}
  \label{fig:case2_pw}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\figmed]{sphx_glr_results_001_2_00x.png}
  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body}
  \label{fig:case2_i}
\end{figure}

The candidate solutions identified by the inversion process produce light curves which adequately approximate the measurements, although they do not reach the lowest observed values. This indicates that the true material is more dominated by specular reflections than our assumed coefficients. The candidates show some rough clustering in both orientation and angular velocity space, although they vary significantly in angular velocity magnitude. The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial guess, possibly indicating that it was reasonably unbiased. Taken together, these results indicate that the object could be spinning in a number of attitude profiles, and that more information is needed about its true mass distribution and reflectivity to further refine the solution set.

\subsection{ECHOSTAR II}

>>>>

\section{Global TODO list}

\begin{itemize}
  \item Add the constant term back into the objective function (in text and code) so we can start treating it as a real likelihood, useful for the future
  \item Mention the unknown solar panel rotation angle for ECHOSTAR
  \item Update perspective on some the case 1a angular velocity plot
  \item Update $n_\text{guess}$ values everywhere to reflect the scripts
  \item Update $\omega$ sampling scheme to go from $0.5-2$ as reflected in scripts
  \item Re-run all cases to ensure that values are up to date with recent changes, update where those statistics are reported in the text
  \item Try to get rid of big gaps where possible
\end{itemize}

@@

If no initial attitude guess is available, as in most cases of passive debris observation, the inversion problem must be solved globally. Filtering approaches can be applied here too; particle \cite{linares2014particle, holzinger2014} and Bayesian Multi-Hypothesis filters \cite{burton2021two, cabrera2023} have shown promise in prior work, but still rely on high-quality priors as with the Kalman filters. In the absence of prior knowledge about the object's attitude state, simulated annealing \cite{gagnon2024, clark2020}, genetic algorithms \cite{gagnon2024, piergentili2017, clark2020}, and particle swarm optimizers (PSOs) \cite{clark2020, clark2022, burton2024journal, burton2024scitech, gagnon2024, gagnon2025} can be applied to search the solution space exhaustively.

@@

In this paper, we describe a global full-state attitude solver inspired by recent PSO work \cite{burton2024journal} but is easier to tune, can be fully parallelized, and is robust to significant measurement noise \cite{robinson2025att}. Further, our method differs from other global estimation methods in that we do not seek to find one optimal solution and instead search for a collection of high-quality local minima to directly study the many ambiguous solutions for a given light curve. Our approach is designed specifically for use with real data as we take time-varying measurement noise and uncertainties in the body moments of inertia into account explicitly.

@@

Our multi-start solution method begins by creating $n_\text{guess}$ guesses scattered throughout the solution space -- possibly sampled from a prior distribution based on available knowledge. In the most basic case, the initial state guess is sampled for guess $i$ at iteration $k=0$:

\begin{equation}
 \vctr{x}_{0,i}(0) = \begin{bmatrix}\vctr{p}_{i,0}(0) \\ \vctr{\omega}_{i,0}(0) \\ \left(J_y / J_x\right)_{i,0} \\ \left(J_z / J_x\right)_{i,0}\end{bmatrix}.
\end{equation}

In this work, initial guesses for the state vector are sampled from a prior distribution, usually defined independently for the MRP, angular velocity, and inertia tensor ratios. If no information is available, a uniform distribution is sampled from for the first six components of the state. In particular, $\vctr{p}_{i,0}(0)$ is uniformly randomly sampled from orientation space, $\vctr{\omega}_{i,0}(0)$ is randomly sampled from a uniform distribution of magnitudes up to $\norm{\vctr{\omega}}_\text{max}$ -- the maximum angular rate at which to initialize guesses. We determine $\norm{\vctr{\omega}}_\text{max}$ based on the maximum power positive frequency in the observed light curve $f^\star$ in Hertz and an arbitrary safety factor $f_\omega \geq 1$, selected as $f_\omega = 1.25$ for this work:

@@

Uniformly sampling over angular velocity directions, we can express the default sampling strategy for an angular velocity guess as:

\begin{equation} \label{eq:omega_sampler}
  \vctr{\omega}_{i,0}(0) = \frac{m}{\sqrt{g_1^2+g_2^2+g_3^2}} \cdot \begin{bmatrix} g_1 & g_2 & g_3 \end{bmatrix}^T,
\end{equation}

\noindent
where $g_{1:3} \in \mathcal{N}(0,1)$ and $m=\mathcal{U}(0,\norm{\vctr{\omega}}_\text{max})$.

@@

Similarly for the MRP, we first sample a quaternion $\vctr{q}_i(0) = [q_x, q_y, q_z, q_w]^T$ uniformly over orientation space:

\begin{equation}
  \vctr{q}_i(0) = \frac{1}{\sqrt{g_1^2 + g_2^2 + g_3^2 + g_4^2}} \begin{bmatrix} g_1 & g_2 & g_3 & g_4 \end{bmatrix}^T,
\end{equation}

\noindent
and project the quaternion into MRP space:

\begin{equation} \label{eq:mrp_sampler}
  \vctr{p}_{i,0}(0) = \frac{1}{1 + q_w} \begin{bmatrix} q_x & q_y & q_z \end{bmatrix}^T.
\end{equation}

Unlike the MRP and angular velocity, the default sampling method for inertia ratios initial guesses $\left(J_y / J_x\right)_{i,0}$ and $\left(J_z / J_x\right)_{i,0}$ draws from independent, identically distributed Gaussian prior distributions with $\sigma_J$ centered at estimated values derived from the known geometry of the object $\left(J_y / J_x\right)_\text{est}$ and $\left(J_z / J_x\right)_\text{est}$, such that:

\begin{equation}
  \begin{bmatrix}
    \left(J_y / J_x\right)_{i,0} \\ \left(J_z / J_x\right)_{i,0}
  \end{bmatrix} \sim \mathcal{N} \left(\begin{bmatrix}
    \left(J_y / J_x\right)_\text{est} \\
    \left(J_z / J_x\right)_\text{est} \end{bmatrix}, \sigma_J I_2 \right),
\end{equation}

\noindent
where $I_2$ is the $2\times2$ identity matrix.

@@

\noindent
producing a converged initial state $\vctr{x}_{k,i}(0)$ which is a good local guess to have produced the observations. In this way, we globally search the solution space up to the density implied by $n$. Using a simpler algorithm like gradient descent can still reach convergence, but will take significantly more objective function evaluations due to the high nonlinearity of the attitude inversion problem.

@@

\begin{equation} \label{eq:ell_selection_criteria}
  e^{-f\left(\vctr{x}_{i,k}(0)\right)} > \ell e^{-f\left(\vctr{x}^\star_{k}(0)\right)}.
\end{equation}

Here, $\vctr{x}^\star_{k}(0) = \argmin_{\vctr{x}_{i,k}(0)} f\left(\vctr{x}_{i,k}(0)\right) $ is the highest likelihood solution found after optimization.

@@

The inversion method in this paper is not dependent on a particular approach for computing light curves and their expected standard deviations. Here, we present the models used to produce the results in Sections \ref{sec:synth_results} and \ref{sec:real_results}.

@@

As the objects we investigate in this work are well-approximated with relatively simple non-convex meshes, we use the semi-analytic method to efficiently compute unshadowed areas via the Sutherland-Hodgeman algorithm \cite{sutherland1974} for each polygon intersection. Further explanation of the polygon clipping procedure for computing the total shadowed area and accelerating this computation during inversion with the so-called shadow cache is included here \textbf{TODO: should this be a `here'? wording this felt weird if we're avoiding saying like `the authors'"} \cite{robinson2025att}.

@@

\section{Attitude Inversion Results for Synthetic Data} \label{sec:synth_results}

To demonstrate the effectiveness of our inversion method, we first present results for a simulated light curve where the ground truth attitude profile is available for comparison.

@@

In order to make this simulated case realistic, we use the presented measurement noise models and introduce a mismatch between the ground truth object used to generate the measurements and the model used in inversion. We produce a noisy, observed light curve for a high-fidelity model of the Saturn V second stage scaled down by a factor $10/3$ to act as a template rocket body object, shown on the right of Figure \ref{fig:rb_models}.

@@

\begin{table}[H]
  \centering
  \caption{Reflection properties of materials used in the synthetic data test case. The white paint parameters are estimated qualitatively}
  \vspace*{6pt}
  \begin{tabular}{|l|l|l|l|}
  \hline
  \textbf{Material} & $C_d$ & $C_s$ & $n$ \\ \hline
 Aluminum \cite{fankhauser2023} & $0.34$ & $0.40$ & $8.9$ \\ \hline
 White paint & $0.9$ & $0.1$ & $1$ \\ \hline
  \end{tabular}
  \label{tb:synth_matprops}
\end{table}

@@

\begin{table}[H]
  \centering
  \caption{True object attitude profile for all test cases}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
 Initial $\vctr{p}(0)$ & $-\frac{1}{3} \begin{bmatrix} 1 & 1 & 1 \end{bmatrix}^T$ \\ \hline
 True $\vctr{\omega}(0)$ ($\text{rad}\cdot\text{s}^{-1}$) & $\vctr{\omega}(0) = \begin{bmatrix} 0.03 & 0.06 & 0.03 \end{bmatrix}^T$ \\ \hline
 Inertia tensor ratios & $J_y / J_x = 1, \: J_z / J_x = 0.25$ \\ \hline
 External torque $\vctr{M}$ & $\begin{bmatrix} 0 & 0 & 0 \end{bmatrix}^T$ \\ \hline
  \end{tabular}
  \label{tb:synth_att}
\end{table}
\FloatBarrier

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\figmed]{light_curves_mag_rocket_body_sdc.png}
  \caption{Simulated mean and sampled observations for the rocket body in black in V-band magnitude, with the variance included in blue.}
  \label{fig:obs_mag_synth}
\end{figure}

@@

\subsection{Case 1a and 1b: Synthetic Light Curve Optimization Results}

@@

Two distinct inversion cases are computed. Case 1a presents inversion results given no attitude information besides an estimate of the inertia tensor ratios derived from the known geometry, while Case 1b presents results given good initial guesses for orientation and angular velocity. Tables \ref{tb:case1a_prior} and \ref{tb:case1b_prior} summarize the initial condition sampling strategies employed for each case.

For both cases, a total of $n_\text{guess}=10^5$ initial state guesses are created via the sampling schemes described their respective tables. BFGS minimization is then performed for each initial guess to locally solve the optimization problem given by Equation \ref{eq:opt_problem}.

\begin{table}[H]
  \centering
  \caption{Initial condition sampling parameters used for inversion in Case 1a}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  $\vctr{p}$ sampling & Uniform, via Equation \ref{eq:mrp_sampler} \\ \hline
  $\vctr{\omega}$ sampling & Uniform, via Equation \ref{eq:omega_sampler} \\ \hline
  $J$ sampling & Gaussian, via Equation \ref{eq:omega_sampler} \\ \hline
  $\sigma_J$ & $0.05$ $[\text{nondim}]$ \\ \hline
  $\norm{\vctr{\omega}}_\text{max}$ & $5.26$ $[\text{deg}\cdot \text{s}^{-1}]$ \\ \hline
\end{tabular}
  \label{tb:case1a_prior}
\end{table}

\begin{table}[H]
  \centering
  \caption{Informative prior parameters used for inversion in Case 1b. The standard deviations $\sigma_\omega$ and $\sigma_J$ are for spherically-symmetric Gaussians with centered on the ground truth. MRPs are sampled by rotating by $\theta\sim \mathcal{N}(0,\sigma_p)$ degrees away from the ground truth orientation around an axis uniformly sampled on the sphere.}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
  $\vctr{p}$ sampling & Gaussian, $\sigma_\omega = 0.05$ $[\text{deg}\cdot\text{s}^{-1}]$ \\ \hline
  $\vctr{\omega}$ sampling & Gaussian, $\sigma_p=20$ $[\text{deg}]$ \\ \hline
  $J$ sampling & Gaussian, $\sigma_J = 0.05$ $[\text{nondim}]$ \\ \hline
\end{tabular}
  \label{tb:case1b_prior}
\end{table}

Solutions are then classified as candidates for the ground truth with $\ell = 1/10$ via Equation \ref{eq:ell_selection_criteria}. Figure \ref{fig:conv_lcs_synth} shows the mean light curves produced by the identified candidate solutions. There is good agreement between the candidate light curves and the observations, indicating that high-quality local minima have been identified.

\begin{figure}[H]
  \centering
  \includegraphics[width=\figbig]{converged_solutions_1_Rocket Body.png}
  \caption{Candidate solution light curves found through BFGS optimization.}
  \label{fig:conv_lcs_synth}
\end{figure}

Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} display the candidate MRP and angular velocity of the candidate solutions determined via the procedure in Section \ref{sec:candidate_sols}. In the left Case 1a figures for orientation and angular velocity, we notice that the ground truth lies squarely on a surface containing numerous candidate solution. This indicates that the local minima have correctly converged to the non-singular subspace of states that approximate the observations well for the assumed shape model. In angular velocity space, shown in Figure \ref{fig:omega_sols1}, most solutions fall onto the surface of two open-ended cylinders -- one with half the radius of the larger cylinder containing the truth. Similarly in orientation space, shown in Figure \ref{fig:sigma_sols1}, the solutions fall on two distinct manifolds, one containing the truth -- likely resulting from a reflection of the object's body frame about the bisector between the Sun and observer vectors \cite{marto2024, burton2024journal}. These additional families of solutions highlight the reality that a wide variety of attitude profiles may produce nearly identical observations.

The Case 1b results on the right within Figures \ref{fig:sigma_sols1} and \ref{fig:omega_sols1} show that the majority of the initial guesses sampled close to the truth have converged to a local subset of the solution manifolds identified by Case 1a. We should not expect these solutions to have shrunk down exponentially towards the ground truth -- the mismatch in the assumed shape model and the measurement noise inherently limits the minimum size of the feasible solution space.

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sigma_solsRocket Body.png}
    \caption{}
    \label{fig:sigma_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sigma_solsRocket Body Prior.png}
    \caption{}
    \label{fig:sigma_sols1b}
  \end{subfigure}

  \caption{Candidate solution initial orientation MRP vectors for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.}
  \label{fig:sigma_sols1}
\end{figure}

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{omega_sols_Rocket Body.png}
    \caption{}
    \label{fig:omega_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{omega_sols_Rocket Body Prior.png}
    \caption{}
    \label{fig:omega_sols1b}
  \end{subfigure}
  \caption{Candidate solution initial body-frame angular velocity vectors for Cases 1a (left) and 1b (right), with the ground truth highlighted in red. TODO: change the perspective on the angular velocities}
  \label{fig:omega_sols1}
\end{figure}

Figure \ref{fig:i_sols1} displays the inertia tensor solutions for Cases 1a and 1b. Notably, the majority of candidate solutions remain within a few standard deviations of the mean of their initial distribution, with the remaining minority exploring directions that maintain the same precession rate. When a good initial guess for the other attitude state components is available in Case 1b, the inertia tensor solutions are more well-clustered, although due to the low observability of the rocket body's spin rate in the observed light curve, there are still ambiguous solutions with higher or lower $J_z / J_x$ values.

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body.png}
    \caption{}
    \label{fig:i_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_2_Rocket Body Prior.png}
    \caption{}
    \label{fig:i_sols1b}
  \end{subfigure}
  \caption{Candidate solution inertia ratios for Cases 1a (left) and 1b (right), with the ground truth highlighted in red.}
  \label{fig:i_sols1}
\end{figure}

Figure \ref{fig:w_vs_ang_error_sols1} summarizes the distribution of candidate solutions with respect to the ground truth by displaying the angular velocity magnitude error and angular orientation error compared to the ground truth. Case 1a shows significant variability in angular velocity magnitude, although the mean error is low at $2\%$. Due to the underdetermined orientation solution space, there are far more solutions at high angular offsets from the truth orientation, leading to a large mean angular error of $127^\circ$. For Case 1b, the mean angular error is much lower at $25^\circ$, although convergence here is still inherently limited by the noisy measurements and model mismatch.

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body.png}
    \caption{}
    \label{fig:w_vs_ang_error_sols1a}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{converged_solutions_3_Rocket Body Prior.png}
    \caption{}
    \label{fig:w_vs_ang_error_sols1b}
  \end{subfigure}

  \caption{Candidate solution angular orientation error and angular velocity magnitudes for Cases 1a (left) and 1b (right) compared to the ground truth}
  \label{fig:w_vs_ang_error_sols1}
\end{figure}

@@

\section{Attitude Inversion Results for Real Data} \label{sec:real_results}

@@

\subsection{Case 2: Delta I Upper Stage}

In Case 2, we perform attitude inversion on real data observed for a Delta I upper stage rocket body. The chosen Delta I upper stage (COSPAR ID 1984-115C) is a Star-37E \cite{delta3914_astronautix} solid fuel apogee kick motor, displayed in Figure \ref{fig:star37e}. It is $d=0.93$ meters in diameter and is approximately $h=1.7$ meters in length \cite{star37e_astronautix, star37_gunter}.

\begin{figure}[H]
  \centering
  \includegraphics[width=0.25\textwidth]{star37e.JPG}
  \caption{Star-37E upper stage \cite{star37_af}.}
  \label{fig:star37e}
\end{figure}

@@

This light curve shows clear periodicity, but has a high measurement standard deviation and a relatively low sampling rate due to the object's low brightness and the sensor readout time. Inversion is performed using $n_\text{guess}=10^4$ initial guesses with the material and angular velocity assumptions listed in Table \ref{tb:case2_ass}. As in Case 1a, orientation and angular velocity initial guesses are sampled via Equations \ref{eq:mrp_sampler} and \ref{eq:omega_sampler}, respectively. All initial inertia tensor guesses are set to the estimated $J_\text{R/B}$. The candidate solution results are presented in terms of their mean light curves in Figure \ref{fig:case2_s}, orientation MRPs and angular velocities in Figure \ref{fig:case2_pw}, and inertia ratios in Figure \ref{fig:case2_i}.

\begin{table}[H]
  \centering
  \caption{Delta I rocket body inversion assumptions}
  \vspace*{6pt}
  \begin{tabular}{|l|l|}
  \hline
  \textbf{Variable} & \textbf{Value} \\ \hline
 Uniform BRDF & $C_d=0.2$, $C_s=0.8$, $n=30$ \\ \hline
 $\norm{\vctr{\omega}}_\text{max}$ & $8.31$ $\text{deg} \cdot \text{s}^{-1}$ \\ \hline
 $\ell$ & $1/100$ \\ \hline
\end{tabular}
  \label{tb:case2_ass}
\end{table}

\begin{figure}[H]
  \centering
  \includegraphics[width=\figbig]{sphx_glr_results_002_2_00x.png}
  \caption{Candidate solution light curves compared to the real measurements in ADU for the Delta I (Star-37E) rocket body.}
  \label{fig:case2_s}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sphx_glr_results_006.png}
    \caption{}
    \label{fig:case2_pwa}
  \end{subfigure}
  \hfill
  \begin{subfigure}[t]{0.23\textwidth}
    \centering
    \includegraphics[width=\textwidth]{sphx_glr_results_005.png}
    \caption{}
    \label{fig:case2_pwb}
  \end{subfigure}

  \caption{Candidate solution orientation MRPs (left) and angular velocity vectors (right) for the Delta I (Star-37E) rocket body}
  \label{fig:case2_pw}
\end{figure}

@@

\begin{figure}[H]
  \centering
  \includegraphics[width=\figsmall]{sphx_glr_results_001_2_00x.png}
  \caption{Candidate solution inertia ratio distribution for the Delta I (Star-37E) rocket body}
  \label{fig:case2_i}
\end{figure}

The candidate solutions identified by the inversion process produce light curves which adequately approximate the measurements, although they do not reach the lowest observed values. This indicates that the true material is more dominated by specular reflections than our assumed coefficients. The candidates show some rough clustering in both orientation and angular velocity space, although they vary significantly in angular velocity magnitude. The inertia tensor solutions display similar variability, although the solutions are spread relatively uniformly from the initial guess, possibly indicating that the geometry-driven estimate is approximately unbiased. Taken together, these results indicate that the object could be spinning in a number of attitude profiles, although the light curve inversion process has significantly narrowed the search space to small subsets of the original unconstrained eight-dimensional volume. These constraints may be combined with additional light curves, more information about the object's true mass distribution, or reflectivity to further refine the solution set.

\subsection{Case 3: ECHOSTAR II}
